---
title: "CourseProject1"
author: "James"
date: "17/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE}
# Load packages 
library(tidyverse)
library(ggplot2)

# Load dataset

activity <- read.csv(".\\activity.csv")
# StormData <- read.csv("repdata_data_StormData1.csv")

# glimpse(activity)
```
```{r, echo=FALSE}

activity <- read.csv(".\\activity.csv")


#glimpse(activity)

# Convert data column to date data type
activity$date <- as.Date(activity$date)
#glimpse(activity)



df <- activity

# summarise steps taken by date and interval

df <- df %>% group_by(date, interval) %>% summarise(steps = sum(steps, na.rm = T))


```

## What is mean total number of steps taken per day?


```{r, echo=FALSE}
## Number of steps per day

df_day <- df %>% group_by(date) %>% summarise(steps = sum(steps, na.rm = T))
df_day_mean <- df %>% group_by(date) %>% summarise(steps = mean(steps, na.rm = T))

#df_day
df_day_mean

```

## Make a histogram of the total number of steps taken each day

```{r echo=FALSE, fig.align='center', fig.height= 5, fig.width= 10}

hist(df_day$steps)

```

## Calculate and report the mean and median of the total number of steps taken per day

```{r echo=FALSE}

df_mean_median <- df %>% group_by(date) %>% summarise(Mean_steps = mean(steps, na.rm = T),Median_steps = median(steps, na.rm = T))

df_mean_median
# 
# 
# mean_Steps <- mean(activity$steps[!is.na(activity$steps)])
# mean_Steps
# 
# # summary(activity$steps[!is.na(activity$steps)])
# median_Steps <- median(activity$steps[!is.na(activity$steps)])
# median_Steps
```


## What is the average daily activity pattern?

```{r echo=FALSE}

# glimpse(activity)

mean_Steps <- mean(df_day$steps)
mean_Steps

# 
# Avg_daily <- activity %>% group_by(date, interval) %>% summarise(Steps = sum(steps, na.rm = T))
# 
# 
# 
# plot(Avg_daily$interval, steps, type = "l")

```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=FALSE}

# glimpse(activity)
df_Interval <- df %>% group_by(interval) %>% summarise(total = steps) %>% arrange(-total)

Max <- max(df_Interval$total)
Max_Interval <- df_Interval$interval[df_Interval$total==max(df_Interval$total)]

Max
Max_Interval

```

5 minute interval number with maximum number `r Max`  is interval number `r Max_Interval`

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with 

```{r echo=FALSE}
missingData <- sum(is.na(activity$steps))
missingData
```

There are a total of `r missingData` missing data in activity dataset



## Devise a strategy for filling in all of the missing values in the dataset.

```{r echo=FALSE}

activity_cleaned <- activity
activity_cleaned[is.na(activity_cleaned$steps), "steps"] <- mean_Steps 

```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r echo=FALSE}

hist(activity_cleaned$steps)
```


```{r echo=FALSE}
mean_Steps_clean <- mean(activity_cleaned$steps[!is.na(activity_cleaned$steps)])
#mean_Steps_clean

# summary(activity_cleaned$steps[!is.na(activity_cleaned$steps)])
median_Steps_clean <- median(activity_cleaned$steps[!is.na(activity_cleaned$steps)])
#median_Steps_clean

paste("Mean   = ", as.character(round(mean_Steps_clean),0), sep = "")
paste("Median = ", as.character(round(median_Steps_clean),0), sep = "")

```

This shows imputing missing data using mean does not affect the median total daily number of steps but the mean value.


## Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r echo=FALSE}

# head(df)

activity_updated <- activity %>% 
  select(date, interval, steps) %>% 
  mutate(weekday = weekdays(activity$date, abbreviate = T)) %>% 
  mutate(weekday1 = 
           case_when(    
              weekday == "Mon" ~ "weekday",
              weekday == "Tue" ~ "weekday",
              weekday == "Wed" ~ "weekday",
              weekday == "Thu" ~ "weekday",
              weekday == "Fri" ~ "weekday",
              TRUE ~ "weekend")) %>% 
  mutate(weekday1 = as.factor(weekday1))

#head(activity_updated)
```

## Make a panel plot containing a time series plot of the 5-minute interval (x-axis)

```{r echo=FALSE}

#plot(wt,disp, main="Scatterplot of wt vs disp")

activity_updated1 <- activity_updated %>% group_by(interval, weekday1) %>% summarise(mean_step = mean(steps, na.rm = T))

activity_updated1 %>% ggplot(aes(interval, mean_step)) +
  geom_line() +
  facet_grid(weekday1 ~.)
```















